---
title: "Group 8 Project ITEC-620"
author: "Loc Le"
date: "2023-11-30"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages
```{r}
library(tree)
library(tidyverse)
library(dplyr)
library(class)
library(fastDummies)
library(FNN)
```

**IMPORTANT**

- Use specific Amazon.csv file uploaded in OneDrive folder
- Run each code chunk one by one (do not run all)

Loading Kaggle Dataset (Saved as Amazon.csv)
```{r}
Amazon <- read.csv("Amazon.csv")
```

Formatting Dataset
```{r}
# ONLY RUN THIS CODE CHUNK ONCE

Amazon <- Amazon[,-1] # Remove Timestamp Column
```

```{r}
# Creates a table of each variable to check for weird values such as N/As, blanks, etc. and infrequently occuring values that can be lumped into an "Other" category per the professor's suggestion

variable_names <- names(Amazon)

frequency_tables <- list()

for (variable in variable_names) {
  frequency_tables[[variable]] <- table(Amazon[[variable]])
}

for (i in seq_along(frequency_tables)) {
  cat("Variable:", names(frequency_tables)[i], "\n")
  print(frequency_tables[[i]])
  cat("\n")
}
```

```{r}
# Removing rows with weird values

Amazon <- Amazon[-c(23,120,383),]
```

```{r}
# Gender

Amazon <- Amazon %>% 
  mutate(Gender = str_replace(Gender, 
                                           "Prefer not to say", "Others"))
```

```{r}
# Lumping purchase subcategories into their parent categories to reduce total number of categories

# Beauty and Personal Care

Amazon <- Amazon %>% 
  mutate(Purchase_Categories = str_replace(Purchase_Categories, 
                                           "Beauty and Personal Care;Clothing and Fashion", "Beauty and Personal Care"))

Amazon <- Amazon %>% 
  mutate(Purchase_Categories = str_replace(Purchase_Categories, 
                                           "Beauty and Personal Care;Home and Kitchen", "Beauty and Personal Care"))

Amazon <- Amazon %>% 
  mutate(Purchase_Categories = str_replace(Purchase_Categories, 
                                           "Beauty and Personal Care;others", "Beauty and Personal Care"))

# Clothing and Fashion

Amazon <- Amazon %>% 
  mutate(Purchase_Categories = str_replace(Purchase_Categories, 
                                           "Clothing and Fashion;Home and Kitchen", "Clothing and Fashion"))

Amazon <- Amazon %>% 
  mutate(Purchase_Categories = str_replace(Purchase_Categories, 
                                           "Clothing and Fashion;others", "Clothing and Fashion"))

# Groceries and Gourmet Food

Amazon <- Amazon %>% 
  mutate(Purchase_Categories = str_replace(Purchase_Categories, 
                                           "Groceries and Gourmet Food;Beauty and Personal Care", "Groceries and Gourmet Food"))

Amazon <- Amazon %>% 
  mutate(Purchase_Categories = str_replace(Purchase_Categories, 
                                           "Groceries and Gourmet Food;Clothing and Fashion", "Groceries and Gourmet Food"))

Amazon <- Amazon %>% 
  mutate(Purchase_Categories = str_replace(Purchase_Categories, 
                                           "Groceries and Gourmet Food;Home and Kitchen", "Groceries and Gourmet Food"))

Amazon <- Amazon %>% 
  mutate(Purchase_Categories = str_replace(Purchase_Categories, 
                                           "Groceries and Gourmet Food;Home and Kitchen;others", "Groceries and Gourmet Food"))

Amazon <- Amazon %>% 
  mutate(Purchase_Categories = str_replace(Purchase_Categories, 
                                           "Groceries and Gourmet Food;others", "Groceries and Gourmet Food"))

# Home and Kitchen

Amazon <- Amazon %>% 
  mutate(Purchase_Categories = str_replace(Purchase_Categories, 
                                           "Home and Kitchen;others", "Home and Kitchen"))
```

```{r}
# Lumping infrequent values in Service_Appreciation to an "Other" category

# All the above

Amazon <- Amazon %>% 
  mutate(Service_Appreciation = str_replace(Service_Appreciation, 
                                           "All the above", "Other"))

# Customer service

Amazon <- Amazon %>% 
  mutate(Service_Appreciation = str_replace(Service_Appreciation, 
                                           "Customer service", "Other"))

# Quick delivery

Amazon <- Amazon %>% 
  mutate(Service_Appreciation = str_replace(Service_Appreciation, 
                                           "Quick delivery", "Other"))
```


```{r}
# Lumping similar values together as well as infrequent values in Improvement_Areas to an "Other" category

# Other

Amazon <- Amazon %>% 
  mutate(Improvement_Areas = str_replace(Improvement_Areas, 
                                           "Add more familiar brands to the list", "Other"))

Amazon <- Amazon %>% 
  mutate(Improvement_Areas = str_replace(Improvement_Areas, 
                                           "better app interface and lower shipping charges", "Other"))

Amazon <- Amazon %>% 
  mutate(Improvement_Areas = str_replace(Improvement_Areas, 
                                           "Irrelevant product suggestions", "Other"))

Amazon <- Amazon %>% 
  mutate(Improvement_Areas = str_replace(Improvement_Areas, 
                                           "I don't have any problem with Amazon", "Other"))

Amazon <- Amazon %>% 
  mutate(Improvement_Areas = str_replace(Improvement_Areas, 
                                           "I have no problem with Amazon yet but others tell me about the refund issues", "Other"))
Amazon <- Amazon %>% 
  mutate(Improvement_Areas = str_replace(Improvement_Areas, 
                                           "Nil", "Other"))

Amazon <- Amazon %>% 
  mutate(Improvement_Areas = str_replace(Improvement_Areas, 
                                           "No problems with Amazon", "Other"))

Amazon <- Amazon %>% 
  mutate(Improvement_Areas = str_replace(Improvement_Areas, 
                                           "Nothing", "Other"))

Amazon <- Amazon %>% 
  mutate(Improvement_Areas = str_replace(Improvement_Areas, 
                                           "Scrolling option would be much better than going to next page", "Other"))

Amazon <- Amazon %>% 
  mutate(Improvement_Areas = str_replace(Improvement_Areas, 
                                           "User interface", "Other"))

Amazon <- Amazon %>% 
  mutate(Improvement_Areas = str_replace(Improvement_Areas, 
                                           "App UI", "Other"))

# Product quality and accuracy

Amazon <- Amazon %>% 
  mutate(Improvement_Areas = str_replace(Improvement_Areas, 
                                           "Quality of product is very poor according to the big offers", "Product quality and accuracy"))
```

```{r}
# Creates a new column called Purchase_Frequency_Num for transforming Purchase_Frequency values into numeric values per the professor's suggestion

Amazon$Purchase_Frequency_Num <- 0.5 # Less than once a month

Amazon$Purchase_Frequency_Num[Amazon$Purchase_Frequency == "Once a month"] <- 1

Amazon$Purchase_Frequency_Num[Amazon$Purchase_Frequency == "Few times a month"] <- 3

Amazon$Purchase_Frequency_Num[Amazon$Purchase_Frequency == "Once a week"] <- 5

Amazon$Purchase_Frequency_Num[Amazon$Purchase_Frequency == "Multiple times a week"] <- 7
```

Less than once a month = 0.5

Once a month = 1

Few times a month = 3

Once a week = 5

Multiple times a week = 7

Exploratory Data Analysis:

```{r}
ggplot(data = Amazon, mapping = aes(x = Purchase_Frequency, fill = Gender)) +
  geom_bar(position = "dodge")
```
Majority of Amazon customers shop there a few times a month and females are the most prominent shoppers.

```{r}
ggplot(data = Amazon, mapping = aes(x = Purchase_Categories, fill = Gender)) +
  geom_bar(position = "dodge")
```
The most popular category at Amazon is Beauty and Personal Care and it is where the majority of females are shopping.

```{r}
Amazon$Age_Category <- ifelse(Amazon$age <= 20, "1-20",
                              ifelse(Amazon$age <= 40, "21-40",
                                     ifelse(Amazon$age <= 60, "41-60", "61+")))
```

```{r}
ggplot(data = Amazon, mapping = aes(x = Age_Category, fill = Gender)) +
  geom_bar(position = "dodge")
```
The majority of female shoppers are between the ages of 21-40



---

```{r}
# ONLY RUN THIS CODE CHUNK ONCE

Amazon <- Amazon[,-3] # Remove old Purchase_Frequency column
```

```{r}
# ONLY RUN THIS CODE CHUNK ONCE

Amazon <- Amazon[,-23] # Remove Age_Category column
```

```{r}
# Creating dummy variables from categorical variables

Amazon <- dummy_cols(Amazon, select_columns=c("Gender", "Purchase_Categories", "Personalized_Recommendation_Frequency", "Browsing_Frequency", "Product_Search_Method", "Search_Result_Exploration", "Add_to_Cart_Browsing", "Cart_Completion_Frequency", "Cart_Abandonment_Factors", "Saveforlater_Frequency", "Review_Left", "Review_Reliability", "Review_Helpfulness", "Recommendation_Helpfulness", "Service_Appreciation", "Improvement_Areas"))
```

```{r}
# Remove old categorical variables
Amazon <- Amazon[,-c(2:7,9:15,17,20,21)]
```

```{r}
glimpse(Amazon)
```

```{r}
# Replace spaces in variables with underscores
colnames(Amazon) <- gsub(" ", "_", colnames(Amazon))
```

```{r}
# Manual fix for this variable
names(Amazon)[names(Amazon) == "Service_Appreciation_User-friendly_website/app_interface"] <- "Service_Appreciation_User_friendly_website_or_app_interface"
```

# Regression Analysis

```{r}
set.seed(12345)
training <- sample(1:nrow(Amazon), 0.6*nrow(Amazon))
ycol <- match("Purchase_Frequency_Num",colnames(Amazon))
Amazon.training <- Amazon[training,-ycol]
Amazon.training.results <- Amazon[training,ycol]
Amazon.test <- Amazon[-training,-ycol]
Amazon.test.results <- Amazon[-training,ycol]
```

```{r}
# Linear Regression with all variables (removed variables are baselines)
Amazon.reg.full <- lm(Purchase_Frequency_Num ~ . 
                 -Gender_Male 
                 -Purchase_Categories_Beauty_and_Personal_Care 
                 - Personalized_Recommendation_Frequency_No
                 - Browsing_Frequency_Multiple_times_a_day
                 - Product_Search_Method_categories
                 - Search_Result_Exploration_Multiple_pages
                 - Add_to_Cart_Browsing_Yes
                 - Cart_Completion_Frequency_Never
                 - Cart_Abandonment_Factors_Found_a_better_price_elsewhere
                 - Saveforlater_Frequency_Never
                 - Review_Left_Yes
                 - Review_Reliability_Never
                 - Review_Helpfulness_Yes
                 - Recommendation_Helpfulness_No
                 - Service_Appreciation_Wide_product_selection
                 - Improvement_Areas_Reducing_packaging_waste
                 ,data = Amazon[training,])

Amazon.reg.full.predictions <- predict(Amazon.reg.full,Amazon)[-training]
(mean((Amazon.test.results-Amazon.reg.full.predictions)^2))^0.5
```

RMSE of linear regression model with all variables = 1.990002

```{r}
summary(Amazon.reg.full)
```


```{r}
# New model with insignificant variables removed

Amazon.reg.reduced <- lm(Purchase_Frequency_Num ~ . 
                 -Gender_Male 
                 -Purchase_Categories_Beauty_and_Personal_Care 
                 -Personalized_Recommendation_Frequency_No
                 -Browsing_Frequency_Multiple_times_a_day
                 -Product_Search_Method_categories
                 -Search_Result_Exploration_Multiple_pages
                 -Add_to_Cart_Browsing_Yes
                 -Cart_Completion_Frequency_Never
                 -Cart_Abandonment_Factors_Found_a_better_price_elsewhere
                 -Saveforlater_Frequency_Never
                 -Review_Left_Yes
                 -Review_Reliability_Never
                 -Review_Helpfulness_Yes
                 -Recommendation_Helpfulness_No
                 -Service_Appreciation_Wide_product_selection
                 -Improvement_Areas_Reducing_packaging_waste
                 -age
                 -Customer_Reviews_Importance
                 -Personalized_Recommendation_Frequency.1
                 -Rating_Accuracy
                 -Shopping_Satisfaction
                 -Gender_Others
                 -Personalized_Recommendation_Frequency_Sometimes
                 -Browsing_Frequency_Few_times_a_month
                 -Browsing_Frequency_Few_times_a_week
                 -Browsing_Frequency_Rarely
                 -Product_Search_Method_Filter
                 -Product_Search_Method_Keyword
                 -Product_Search_Method_others
                 -Search_Result_Exploration_First_page
                 -Add_to_Cart_Browsing_Maybe
                 -Cart_Completion_Frequency_Often
                 -Cart_Completion_Frequency_Rarely
                 -Cart_Completion_Frequency_Sometimes
                 -Cart_Abandonment_Factors_Changed_my_mind_or_no_longer_need_the_item
                 -Cart_Abandonment_Factors_High_shipping_costs
                 -Cart_Abandonment_Factors_others
                 -Saveforlater_Frequency_Always
                 -Saveforlater_Frequency_Often
                 -Saveforlater_Frequency_Rarely
                 -Saveforlater_Frequency_Sometimes
                 -Review_Reliability_Heavily
                 -Review_Reliability_Moderately
                 -Review_Reliability_Occasionally
                 -Review_Reliability_Rarely
                 -Review_Helpfulness_No
                 -Review_Helpfulness_Sometimes
                 -Recommendation_Helpfulness_Sometimes
                 -Recommendation_Helpfulness_Yes
                 -Service_Appreciation_Competitive_prices
                 -Service_Appreciation_Other
                 -Service_Appreciation_Product_recommendations
                 -Service_Appreciation_User_friendly_website_or_app_interface
                 -Improvement_Areas_Customer_service_responsiveness
                 -Improvement_Areas_Other
                 -Improvement_Areas_Product_quality_and_accuracy
                 , data = Amazon[training, ])


Amazon.reg.reduced.predictions <- predict(Amazon.reg.reduced,Amazon)[-training]
(mean((Amazon.test.results-Amazon.reg.reduced.predictions)^2))^0.5
```

RMSE of linear regression model with removed variables = 1.979471

```{r}
summary(Amazon.reg.reduced)
```

```{r}
# New model with insignificant variables removed (Iteration 2)

Amazon.reg.reduced2 <- lm(Purchase_Frequency_Num ~ . 
                 -Gender_Male 
                 -Purchase_Categories_Beauty_and_Personal_Care 
                 -Personalized_Recommendation_Frequency_No
                 -Browsing_Frequency_Multiple_times_a_day
                 -Product_Search_Method_categories
                 -Search_Result_Exploration_Multiple_pages
                 -Add_to_Cart_Browsing_Yes
                 -Cart_Completion_Frequency_Never
                 -Cart_Abandonment_Factors_Found_a_better_price_elsewhere
                 -Saveforlater_Frequency_Never
                 -Review_Left_Yes
                 -Review_Reliability_Never
                 -Review_Helpfulness_Yes
                 -Recommendation_Helpfulness_No
                 -Service_Appreciation_Wide_product_selection
                 -Improvement_Areas_Reducing_packaging_waste
                 -age
                 -Customer_Reviews_Importance
                 -Personalized_Recommendation_Frequency.1
                 -Rating_Accuracy
                 -Shopping_Satisfaction
                 -Gender_Others
                 -Personalized_Recommendation_Frequency_Sometimes
                 -Browsing_Frequency_Few_times_a_month
                 -Browsing_Frequency_Few_times_a_week
                 -Browsing_Frequency_Rarely
                 -Product_Search_Method_Filter
                 -Product_Search_Method_Keyword
                 -Product_Search_Method_others
                 -Search_Result_Exploration_First_page
                 -Add_to_Cart_Browsing_Maybe
                 -Cart_Completion_Frequency_Often
                 -Cart_Completion_Frequency_Rarely
                 -Cart_Completion_Frequency_Sometimes
                 -Cart_Abandonment_Factors_Changed_my_mind_or_no_longer_need_the_item
                 -Cart_Abandonment_Factors_High_shipping_costs
                 -Cart_Abandonment_Factors_others
                 -Saveforlater_Frequency_Always
                 -Saveforlater_Frequency_Often
                 -Saveforlater_Frequency_Rarely
                 -Saveforlater_Frequency_Sometimes
                 -Review_Reliability_Heavily
                 -Review_Reliability_Moderately
                 -Review_Reliability_Occasionally
                 -Review_Reliability_Rarely
                 -Review_Helpfulness_No
                 -Review_Helpfulness_Sometimes
                 -Recommendation_Helpfulness_Sometimes
                 -Recommendation_Helpfulness_Yes
                 -Service_Appreciation_Competitive_prices
                 -Service_Appreciation_Other
                 -Service_Appreciation_Product_recommendations
                 -Service_Appreciation_User_friendly_website_or_app_interface
                 -Improvement_Areas_Customer_service_responsiveness
                 -Improvement_Areas_Other
                 -Improvement_Areas_Product_quality_and_accuracy
                 -Gender_Female
                 -Purchase_Categories_Groceries_and_Gourmet_Food
                 , data = Amazon[training, ])


Amazon.reg.reduced2.predictions <- predict(Amazon.reg.reduced2,Amazon)[-training]
(mean((Amazon.test.results-Amazon.reg.reduced2.predictions)^2))^0.5
```

RMSE of linear regression model with removed variables (second iteration) = 1.980209

This is higher than previously so it will not be used.

Therefore, Linear Regression Model 2 is the best

```{r}
summary(Amazon.reg.reduced2)
```


KNN

```{r}
# Find optimal value of k
best.k <- -1
RMSE <- -1
best.RMSE <- 99999999
for (i in seq(from=1, to=100, by=2)) {
  set.seed(12345)
  Amazon.knn <- knn.reg(Amazon.training, Amazon.test, Amazon.training.results, k=i)
  RMSE <- (mean((Amazon.knn$pred - Amazon.test.results)^2))^0.5
  if (RMSE < best.RMSE) {
    best.k <- i
    best.RMSE <- RMSE
  }
}
print(paste("The optimal value of k is",best.k,"with a RMSE of",best.RMSE))
```

```{r}
Amazon.knn <- knn.reg(Amazon.training, Amazon.test, Amazon.training.results, k=25)
(mean((Amazon.knn$pred - Amazon.test.results)^2))^0.5
```

Regression Tree:

```{r}
best.mindev <- -1
RMSE <- -1
best.RMSE <- 99999999
for (i in seq(from=0.0005, to=0.05, by=0.0005)) {
  Amazon.tree <- tree(Purchase_Frequency_Num ~ ., data=Amazon[training,], mindev=i)
  Amazon.tree.predictions <- predict(Amazon.tree,Amazon)[-training]
  RMSE <- (mean((Amazon.test.results-Amazon.tree.predictions)^2))^0.5
  if (RMSE < best.RMSE) {
    best.mindev <- i
    best.RMSE <- RMSE
  }
}
print(paste("The optimal value of mindev is",best.mindev,"with a RMSE of",best.RMSE))
```

```{r}
Amazon.tree <- tree(Purchase_Frequency_Num ~., data=Amazon[training,],mindev=0.045)
Amazon.tree.predictions <- predict(Amazon.tree,Amazon)[-training]
(mean((Amazon.test.results-Amazon.tree.predictions)^2))^0.5
```

```{r}
plot(Amazon.tree)
text(Amazon.tree, cex=0.6)
```


